Article title

### 2017-07-28 paper-webserver.tex
CHANGE: rewrite of the traditional wet-labs tech para

% costs of bacterial infections
Foodborne ilnesses are increasingly difficult to track; food is globally sourced and consumers are purchasing more prepared ingrediants \cite{kozak2013foodborne}.
The resulting consumer infections translate into actual costs: in Canada alone, an estimated 4000 hospitalizations and 105 deaths per year are thought to be the result of various pathogens \cite{thomas2015estimates}.
Futhermore, the associated economic costs have been known for decades \cite{todd1989costs}.
% surveillance
Public-health agencies routinely look for emerging threats, of which \textit{Salmonella enterica} and \textit{Escherichia coli} are amongst the two most common \cite{kozak2013foodborne}.
In \textit{Salmonella enterica} \cite{bell2016recent}, \textit{Escherichia coli} \cite{fratamico2016advances}, and others pathogens \cite{ronholm2016navigating}, a variety of bioinformatic approaches are now based on whole-genome sequencing (WGS).

% traditional reference lab techniques
To begin analyzing genomes, researchers designate particular strains of bacterial isolates by identifiers known as subtypes.
Traditional reference laboratory techniques handled subtyping through methods such as serotyping, detecting the presence of specific cell surface antigens as tested by antibody reaction, pulsed-field gel electrophoresis (PFGE), to identify band patterns, known virulence factors (VFs), or antimicrobial resistance genes (AMR genes), or Sanger sequencing, to determine multi-locus sequence types (MLSTs) \cite{ronholm2016navigating}.
In public health agencies, PFGE is used to track outbreaks by comparing band patterns of bacterial isolates \cite{kozak2013foodborne}. 
The original computational methods leveraging WGS returned subtyping results matching those of wet-lab techniques.
Software such as VirulenceFinder \cite{joensen2014real} compares a pre-collected set of reference VF gene sequences against a WGS result.
Similar approaches are applied to predict AMR genes \cite{mcarthur2013comprehensive}, O-type, and H-type \cite{whiteside2016superphy}.

>>>>

% costs of bacterial infections
Foodborne ilnesses are increasingly difficult to track; food is globally sourced and consumers are purchasing more prepared ingrediants \cite{kozak2013foodborne}.
The resulting consumer infections translate into actual costs: in Canada alone, an estimated 4000 hospitalizations and 105 deaths per year are thought to be the result of various pathogens \cite{thomas2015estimates}.
Futhermore, the associated economic costs have been known for decades \cite{todd1989costs}.
% surveillance
Public-health agencies routinely monitor emerging threats, of which \textit{Salmonella enterica} and \textit{Escherichia coli} are amongst the two most common \cite{kozak2013foodborne}.
In \textit{Salmonella enterica} \cite{bell2016recent}, \textit{Escherichia coli} \cite{fratamico2016advances}, and others pathogens \cite{ronholm2016navigating}, a variety of bioinformatic approaches are now based on whole-genome sequencing (WGS).

% traditional reference lab techniques
To begin analyzing genomes, researchers designate particular strains of bacterial isolates using identifiers known as subtypes.
Traditional reference laboratory techniques handled subtyping through methods such as serotyping, pulsed-field gel electrophoresis (PFGE), or Sanger sequencing, amongst others \cite{ronholm2016navigating}.
PFGE is unique amongst the set; restriction enzymes are used to produce DNA fragments visible as specific banding patterns.
Public health agencies use PFGE to track outbreaks by comparing band patterns of bacterial isolates \cite{kozak2013foodborne}.
However, as WGS rarely produces complete circular genomes and, in PFGE, cutting by a restriction enzyme is dependent on protein structure, there is not a bioinformatic-analog to PFGE.
Serotyping uses antibodies to detect the presence of specific cell surface antigens.
Sanger sequencing is historically used to examine differences in housekeeping genes which are required to maintain cell function, an approach known as multi-locus sequence typing (MLST).
Unlike PFGE, both serotyping and MLST have WGS-based computational methods which mimick results of the wet-lab techniques.
Software such as VirulenceFinder \cite{joensen2014real} compares a pre-collected set of reference VF gene sequences against a WGS result.
Similar approaches are applied to predict AMR genes \cite{mcarthur2013comprehensive}, O-type, and H-type \cite{whiteside2016superphy}.
